---
title: "adhoc-datasetanalysis"
author: "Pega CDH Tools"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{adhoc-datasetanalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
fig.width = 7,
fig.height = 5,
fig.align = "center",
comment = "#>"
)

library(ggplot2)
library(colorspace)

theme_set(theme_minimal())
```

# Ad-Hoc Data Analysis using Dataset Exports

When testing and debugging DSM projects involving data flows, it often is very handy to inspect datasets. There
are a couple of out-of-the-box datasets, like `pxInteractionHistory` that represents the data stored in IH, or 
`pxDecisionResults` that contains the model data and the results of strategies. You can also have custom
datasets created perhaps as the results of running dataflows or when generating data with MonteCarlo datasets.

Any of the datasets that supports the `browse` operation (basically, any that can be read from), can also be
exported. The export format used is multi-line JSON, i.e. every record from the dataset is exported as a single
line containing the JSON data.

Once you have exported a dataset from Pega, you can easily import it and explore it.

In this example we take a look at a sample of `pxInteractionHistory` but this really works for any dataset.

In Pega, export the dataset and download it to your local filesystem. 

```{r, eval=FALSE}
library(cdhtools)

# include below when developing the library
sapply(list.files("~/Documents/pega/cdh-datascientist-tools/r/R", "*.R", full.names = T), source)

ihdata <- readDSExport("Data-pxStrategyResult_pxInteractionHistory", "~/Downloads") 
```

The `cdhtools` package ships with a small sample of IH data

```{r}
data(ihsampledata, package="cdhtools")
knitr::kable(head(ihsampledata, 5))
```

Some of the columns in your dataset may represent date or date-time values. Pega will export these as strings in the
standard Pega format. For analysis it will be convenient to convert them to proper date/time objects so we can easily
report and plot.

```{r}
for (f in c("pxOutcomeTime", "pxDecisionTime")) {
  ihsampledata[[f]] <- cdhtools::fromPRPCDateTime(ihsampledata[[f]])
}
summary(ihsampledata[,grepl("Time",names(ihsampledata)), with=F])
```

With the data in place, you can now do the analyses you want. For the IH data we can for example plot accept rate over time.

BTW, if you're not familiar with the `data.table` syntax below, it would be a good idea to look into this. The internals
of `cdhtools` heavily rely on it also.

```{r, warning=F}
ihsampledata[, period := lubridate::as_date(pxOutcomeTime)]
ihsampledata[, successRate := sum(pyOutcome=="Accepted")/.N, by=c("period","pyChannel")]
ggplot(ihsampledata, aes(period, successRate, color=pyChannel)) + 
  geom_line(size=1) +
  # geom_col(position=position_dodge()) +
  scale_color_discrete_qualitative(name="Channel", guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(limits = c(0, NA), name = "Success Rate", labels = scales::percent) +
  labs(title = "Success Rate over Time", x="Date")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

But you can of course also use a different plot style, for example from the `tidyverse` universe:

```{r, warning=F, message=F}
library(tidyverse)
ggplot(group_by(ihsampledata, pyName, pyChannel, pyIssue) %>% 
         rename(`Channel`=pyChannel) %>% 
         summarise(count = n()) %>% 
         mutate(pct = count/nrow(ihsampledata)), 
       aes(pyName, pct, fill=`Channel`)) + geom_bar(stat="identity") + 
  coord_flip()+ scale_y_continuous(labels = scales::percent)+
  labs(title = "Proposition Distribution", x="Proposition", y="Percentage of Total Offered")
```

