---
title: "adm-explained"
author: "Pega"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{adm-explained}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
# knitr options: https://yihui.name/knitr/options/
knitr::opts_chunk$set(
collapse = TRUE,
fig.width = 7,
fig.height = 5,
fig.align = "center",
comment = "#>"
)

library(ggplot2)
library(scales)

theme_set(theme_minimal())
```

# ADM Model Report Dissected

This notebook shows exactly how all the values in an ADM model report are
calculated. It also shows how the propensity is calculated for a particular
customer.

We use one of the shipped datamart exports for the example. This is a model very similar to one used in some of the ADM PowerPoint/Excel deep dive examples. To load your own data, see the vignette on ADM reporting for examples.

```{r}
library(cdhtools)
library(data.table)

data(admdatamart_models)
data(admdatamart_binning)
```

For the example we use one particular model: FirstMortgage over SMS. You can
use your own data and select a different model.

To explain the ADM model report, we use the "AnnualIncome" predictor as an
example. This too can be swapped for any other predictor when using different
data.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Commented-out chunk useful to find a good example in the data
modelsGoodPreds <- unique((latestSnapshotsOnly(admdatamart_binning) [TotalBins>4 & Type=="numeric" & EntryType == "Active"])$ModelID )
latestSnapshotsOnly(admdatamart_models)[ModelID %in% modelsGoodPreds, c("Positives","Performance","ActivePredictors","Name","Channel")][order(-ActivePredictors)]
print(unique(modelpredictors[TotalBins>4 & Type=="numeric" & EntryType == "Active"]$PredictorName))
```

```{r}
model <- latestSnapshotsOnly(admdatamart_models) [Name == "AutoNew36Months" & Channel == "SMS"]
modelpredictors <- latestSnapshotsOnly(admdatamart_binning)[ModelID == model$ModelID & EntryType != "Inactive"]

predictorbinning <- modelpredictors[PredictorName == "IH.SMS.Outbound.Accepted.pyHistoricalOutcomeCount"][order(BinIndex)]
```

# Model overview

The selected model is shown below. Only the currently active predictors are
used for the propensity calculation, so only showing those.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
modelmetadata <- data.table(Action = paste(model$Issue, model$Group, sep="/"),
                            Channel = model$Channel,
                            Name = model$Name,
                            `Active Predictors` = paste(setdiff(unique(modelpredictors$PredictorName),"Classifier"), collapse = ", "),
                            `Model Performance (AUC)` = model$Performance*100)
kable(t(modelmetadata))  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  column_spec(1, bold=T)
```

## Predictor binning for `r predictorbinning$PredictorName[1]`

The ADM model report will show predictor binning similar to this, with all displayed data coming from fields in the ADM data mart. 
In subsequent sections we'll show how all the data is derived from the number of positives and negatives in each of the bins.

```{r, echo=FALSE}
predictormetadata <- data.table(Name = predictorbinning$PredictorName[1],
                                Range = predictorbinning$Contents[1],
                                Responses = predictorbinning$ResponseCount[1],
                                `# Bins` = predictorbinning$TotalBins[1],
                                `Predictor Performance (AUC)` =  predictorbinning$Performance[1]*100)
kable(t(predictormetadata))  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  column_spec(1, bold=T)
```

```{r, echo=FALSE}
predictorbinning2 <- data.table( `Range/Symbols` = predictorbinning$BinSymbol,
                                 `Responses (%)` = predictorbinning$BinResponseCountPercentage,
                                 `Positives` = predictorbinning$BinPositives,
                                 `Positives (%)` = predictorbinning$BinPositivesPercentage,
                                 `Negatives` = predictorbinning$BinNegatives,
                                 `Negatives (%)` = predictorbinning$BinNegativesPercentage,
                                 # strange that propensity would not be available
                                 `Propensity (%)` = round(predictorbinning$BinPositives/(predictorbinning$BinResponseCount), digits = 4),
                                 `Z-Ratio` = predictorbinning$ZRatio,
                                 `Lift` = predictorbinning$Lift
)
totals <- data.table(`Range/Symbols` = "Grand Total")[, names(predictorbinning2)[2:9] := c(lapply(predictorbinning2[, 2:6], sum), as.numeric(predictorbinning$Positives[1])/predictorbinning$ResponseCount[1], 0.0, 1.0)]

predictorbinning <- predictorbinning2
predictorbinning2 <- rbind(predictorbinning2, totals)

kable(predictorbinning2) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

# Bin statistics

## Positive and Negative ratios

Internally, ADM only keeps track of the total counts of positive and negative responses in each bin. Everything
else is derived from those numbers. The percentages and totals are trivially derived, and the propensity is just 
the number of positives divided by the total. The numbers calculated here match the numbers from the datamart
table exactly.

```{r}
binningDerived <- predictorbinning[, c(1,3,5)] # copy over only the labels, pos and neg counts
binningDerived[, `Responses %` := (Positives+Negatives)/(sum(Positives)+sum(Negatives))]
binningDerived[, `Positives %` := Positives/sum(Positives)]
binningDerived[, `Negatives %` := Negatives/sum(Negatives)]
binningDerived[, Propensity := (Positives)/(Positives+Negatives)]
```

```{r, echo=F}
binningDerived[, `Responses %` := round(100*`Responses %`,2)]
binningDerived[, `Positives %` := round(100*`Positives %`,2)]
binningDerived[, `Negatives %` := round(100*`Negatives %`,2)]
binningDerived[, Propensity := round(Propensity,4)]
kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  column_spec(2:3, bold = T, border_left = T, border_right = T) %>%
  column_spec(4:7, color = "blue") 
binningDerived[, Propensity := (Positives)/(Positives+Negatives)] # put back as we changed it for display purposes
```

## Lift

Lift is the ratio of the propensity in a particular bin over the average propensity. So a value of 1 is the average, larger than 1 means higher propensity, smaller means lower propensity:

```{r}
binningDerived[, Lift := (Positives/(Positives+Negatives)) / (sum(Positives)/sum(Positives+Negatives))]
```

```{r, echo=F}
binningDerived[, `Responses %` := NULL]
binningDerived[, `Positives %` := NULL]
binningDerived[, `Negatives %` := NULL]
binningDerived[, Propensity := NULL]

binningDerived[, Lift := round(Lift,4)]
kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  column_spec(c(2,3), bold = T, border_left = T, border_right = T) %>%
  column_spec(4, color = "blue") 
```

## Z-Ratio

The Z-Ratio is also a measure of the how the propensity in a bin differs from the average, but takes into account the size of the bin and thus is statistically more relevant. It represents the number of standard deviations from the average, so centers around 0. The wider the spread, the better the predictor is.

$$\frac{posFraction-negFraction}{\sqrt(\frac{posFraction*(1-posFraction)}{\sum positives}+\frac{negFraction*(1-negFraction)}{\sum negatives})}$$ 

See also: http://techdocs.rpega.com/display/EPZ/2019/06/21/Z-ratio+calculation+in+ADM.

```{r}
binningDerived[, posFraction := Positives/sum(Positives)]
binningDerived[, negFraction := Negatives/sum(Negatives)]
binningDerived[, `Z-Ratio` := (posFraction-negFraction)/sqrt(posFraction*(1-posFraction)/sum(Positives) + negFraction*(1-negFraction)/sum(Negatives))]
```

```{r, echo=F}
binningDerived[, Lift := NULL]

kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  column_spec(c(2,3), bold = T, border_left = T, border_right = T) %>%
  column_spec(6, color = "blue") 
```

# Predictor AUC

The predictor AUC is the univariate performance of this predictor against the
outcome. This too can be derived from the positives and negatives, e.g. using the *pROC* package.

```{r, warning=F,message=F}
library(pROC)

response = unlist(sapply(1:nrow(predictorbinning),
                         function(r){return(c(rep(T, predictorbinning$Positives[r]), 
                                              rep(F, predictorbinning$Negatives[r])))}))

prediction = unlist(sapply(1:nrow(predictorbinning),
                           function(r){return(rep(predictorbinning$`Propensity (%)`[r], 
                                                  predictorbinning$Positives[r] +
                                                    predictorbinning$Negatives[r]))}))

plot.roc(response, prediction, print.auc=T, col="darkgreen", levels=c(T,F), direction=">")
```

There is also a convenient function in *cdhtools* to calculate it directly from the positives and negatives:

```{r}
cdhtools::auc_from_bincounts(predictorbinning$Positives, predictorbinning$Negatives)
```

# Predictor score and log odds

The score is calculated from the log odds which are simply the ratio of the probabilities of positives and negatives. For the actual calculation in ADM this is modified slightly to avoid division-by-zero problems and is written differently to avoid numeric instability as shown below.

```{r}
binningDerived[, posFraction := Positives/sum(Positives)]
binningDerived[, negFraction := Negatives/sum(Negatives)]
binningDerived[, `Log odds` := log(posFraction/negFraction)]
binningDerived[, `Modified Log odds` := 
                 log(Positives+1/.N) - log(sum(Positives+1)) - 
                 log(Negatives+1/.N) + log(sum(Negatives+1))]
```

```{r, echo=F}
binningDerived[, `Z-Ratio` := NULL]
kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  column_spec(c(2,3), bold = T, border_left = T, border_right = T) %>%
  column_spec(c(6,7), color = "blue") 
```

# Propensity mapping

## Log odds contribution for all the predictors

To get to a propensity, the log odds of the relevant bins of the active predictors are added up and divided by the number of active predictors +1, then used to index in the classifier.

Below an example. From all the active predictors of the model for `r model$name` we pick a value (in the middle for numerics, first symbol for symbolics) and show the (modified) log odds. These log odds values are averaged (added up and divided by number of active predictors + 1), and this is the "score" that is mapped to a propensity value by the classifier (which is constructed using the PAV(A) algorithm).

```{r, echo=F}
binning <- modelpredictors[EntryType=="Active", c("PredictorName", "BinSymbol", "BinIndex", "BinPositives", "BinNegatives", "Type","BinLowerBound", "BinUpperBound"), with=F]
setnames(binning, c("Name", "Value", "Bin", "Positives", "Negatives", "Type", "lobound", "hibound"))

binning[, `Log odds` := log(Positives+1/.N) - log(sum(Positives+1)) - log(Negatives+1/.N) + log(sum(Negatives+1)), by=Name]

binning[,nbins := max(Bin), by=Name]
binning <- binning[Bin == trunc(nbins/2)] # take middle bin
for (r in 1:nrow(binning)) {
  if (binning$Type[r] == "numeric") {
    binning$Value[r] <- trunc((as.numeric(binning$lobound[r]) + as.numeric(binning$hibound[r]))/2) # middle value
  } else {
    binning$Value[r] <- strsplit(binning$Value[r], ",", fixed=T)[[1]][1] # first symbol
  }
}
binning <- binning[, c(1:5,9)]

binning <- rbindlist(list(binning, data.table(Name = "Average Log odds",
                                              Value = "",
                                              `Log odds` = sum(binning$`Log odds`)/(nrow(binning)+1))), use.names = T, fill = T)
kable(binning) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  column_spec(c(1, 2, 3, 4, 5), bold = T, border_left = T, border_right = T) %>%
  column_spec(6, color = "blue") 
```

# Classifier

The success rate is defined as $\frac{positives}{positives+negatives}$ per bin. 

The adjusted propensity that is returned is a small modification (Laplace smoothing) to this and calculated as $\frac{0.5+positives}{1+positives+negatives}$ so empty models return a propensity of 0.5.

```{r, echo=F, warning=F, error=F}
classifier <- modelpredictors[EntryType == "Classifier"][order(BinIndex)]
classifier[, Propensity := BinPositives/(BinPositives+BinNegatives)]
classifier[, AdjustedPropensity := (0.5+BinPositives)/(1+BinPositives+BinNegatives)]

score <- binning$`Log odds`[nrow(binning)] # last one is the average log odds
scorebin <- findInterval(score, as.numeric(classifier$BinLowerBound), left.open = T, all.inside = T)

classifierTable <- userFriendlyADMBinning(classifier)

kable(classifierTable) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  column_spec(c(2,3,4), bold = T, border_left = T, border_right = T) %>%
  column_spec(c(7), color = "blue") 
```


# Final propensity

Below the classifier mapping. On the x-axis the binned scores (log odds values), on the y-axis the Success Rates. Note the returned propensities are following a slightly adjusted formula, see the table above. The bin that contains the calculated score is highlighted.

The score *`r score`* falls in *bin `r scorebin`* of the classifier, so for this set of inputs, the model returns a propensity of `r sprintf("%.2f%%", 100*classifier$AdjustedPropensity[scorebin])`.

```{r, echo=F}
successRateMax <- max(classifier$Propensity, na.rm = T)
if (0 == successRateMax) { successRateMax <- 1 }
secAxisFactor <- ceiling(max(classifier$BinResponseCount)/successRateMax)

  # geom_col(aes(y=(BinPositives+BinNegatives)/secAxisFactor), classifier[BinIndex == scorebin], fill="steelblue3")+
  # geom_col(aes(y=(BinPositives+BinNegatives)/secAxisFactor), classifier[BinIndex != scorebin], fill="grey")+

plotADMBinning(classifier) +
      geom_col(aes(y=BinResponseCount/secAxisFactor), classifier[BinIndex != scorebin], fill="grey")+
      geom_col(aes(y=BinResponseCount/secAxisFactor), classifier[BinIndex == scorebin], fill="steelblue3")+
      geom_line(colour="orange", size=2)+geom_point()+
    geom_hline(data=classifier[1,], mapping = aes(yintercept = Positives/(Positives+Negatives)),
               colour="orange", linetype="dashed")+
  geom_label(data=classifier[scorebin,], 
             mapping=aes(label=sprintf("Returned Propensity: %.2f%%", 100*AdjustedPropensity)), 
             y=classifier[scorebin,]$BinPositives/classifier[scorebin,]$BinResponseCount + 0.0002, 
             vjust=0)
```
