---
title: "adm-details"
author: "Otto Perdeck"
date: "`r Sys.Date()`"
output: 
  pdf_document:
  rmarkdown::html_vignette: 
    toc : yes
vignette: >
  %\VignetteIndexEntry{adm-details}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
  
```{r setup, include = FALSE}
# knitr options: https://yihui.name/knitr/options/
knitr::opts_chunk$set(
collapse = TRUE,
fig.width = 7,
fig.height = 5,
fig.align = "center",
comment = "#>"
)
```

# ADM Model Report Dissected

We will use one of the shipped datamart exports for the example. This is a model very similar to one used in some of the ADM PowerPoint/Excel deep dive examples. To load your own data, see the vignette on ADM reporting for examples.

```{r}
library(cdhtools)
library(data.table)

data(admdatamart_models)
data(admdatamart_binning)

model <- admdatamart_models[pyconfigurationname == "VerySimpleSalesModel" & pyname == "PSDISCOUNT10"][pysnapshottime == max(pysnapshottime)]
modelpredictors <- admdatamart_binning[pymodelid == model$pymodelid]
                                       
predictorname <- "COUNTRY"
predictorbinning <- modelpredictors[pypredictorname == predictorname][order(pybinindex)]
```

The selected model is:

```{r, echo=FALSE}
library(knitr)
library(kableExtra)
modelmetadata <- data.table(Group = paste(model$pyissue, model$pygroup, sep="/"),
                            Name = model$pyname,
                            Predictors = paste(setdiff(unique(modelpredictors$pypredictorname),"Classifier"), collapse = ","),
                            `Model Performance (AUC)` = model$pyperformance*100)
kable(t(modelmetadata))  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  column_spec(1, bold=T)
```

The ADM model report will show predictor binning similar to this, with all displayed data coming from fields in the ADM data mart:

```{r, echo=FALSE}
predictormetadata <- data.table(Name = predictorbinning$pypredictorname[1],
                            Range = predictorbinning$pycontents[1],
                            Responses = predictorbinning$pyresponsecount[1],
                            `# Bins` = predictorbinning$pytotalbins[1],
                            `Predictor Performance (AUC)` =  predictorbinning$pyperformance[1]*100)
kable(t(predictormetadata))  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  column_spec(1, bold=T)
```

```{r, echo=FALSE}
predictorbinning2 <- data.table( `Range/Symbols` = predictorbinning$pybinsymbol,
                       `Responses (%)` = predictorbinning$pybinresponsecountpercentage,
                       `Positives` = predictorbinning$pybinpositives,
                       `Positives (%)` = predictorbinning$pybinpositivespercentage,
                       `Negatives` = predictorbinning$pybinnegatives,
                       `Negatives (%)` = predictorbinning$pybinnegativespercentage,
                       # strange that propensity would not be available
                       `Propensity (%)` = round(predictorbinning$pybinpositives/(predictorbinning$pybinresponsecount), digits = 4),
                       `Z-Ratio` = predictorbinning$pyzratio,
                       `Lift` = predictorbinning$pylift
                       )
totals <- data.table(`Range/Symbols` = "Grand Total")[, names(predictorbinning2)[2:9] := c(lapply(predictorbinning2[, 2:6], sum), as.numeric(predictorbinning$pypositives[1])/predictorbinning$pyresponsecount[1], 0.0, 1.0)]

predictorbinning <- predictorbinning2
predictorbinning2 <- rbind(predictorbinning2, totals)

kable(predictorbinning2) %>%
  kable_styling(bootstrap_options = "striped", full_width = T)
```

# Simple ratios and totals

The counts of positive and negative responses in each bin are the only things that ADM tracks, the rest is derived from these. The percentages and totals are trivially derived, and the propensity is just the number of positives divided by the total.

```{r}
binningDerived <- predictorbinning[, c(1,3,5)] # copy over only the labels, pos and neg counts
binningDerived[, `Responses %` := (Positives+Negatives)/(sum(Positives)+sum(Negatives))]
binningDerived[, `Positives %` := Positives/sum(Positives)]
binningDerived[, `Negatives %` := Negatives/sum(Negatives)]
binningDerived[, Propensity := (Positives)/(Positives+Negatives)]
```

```{r, echo=F}
binningDerived[, `Responses %` := round(100*`Responses %`,2)]
binningDerived[, `Positives %` := round(100*`Positives %`,2)]
binningDerived[, `Negatives %` := round(100*`Negatives %`,2)]
binningDerived[, Propensity := round(Propensity,4)]
kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  column_spec(2:3, bold = T, border_left = T, border_right = T) %>%
  column_spec(4:7, color = "darkblue") 
binningDerived[, Propensity := (Positives)/(Positives+Negatives)] # put back as we changed it for display purposes
```

# Lift

Lift is the ratio of the propensity in a particular bin over the average propensity. So a value of 1 is the average, larger than means higher propensity, smaller means lower propensity.

```{r}
binningDerived[, Lift := (Positives/(Positives+Negatives)) / (sum(Positives)/sum(Positives+Negatives))]
```

```{r, echo=F}
binningDerived[, `Responses %` := NULL]
binningDerived[, `Positives %` := NULL]
binningDerived[, `Negatives %` := NULL]
binningDerived[, Propensity := NULL]

binningDerived[, Lift := round(Lift,4)]
kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  column_spec(c(2,3), bold = T, border_left = T, border_right = T) %>%
    column_spec(4, color = "darkblue") 
```

# Z-Ratio

The Z-Ratio is also a measure of the how the propensity in a bin differs from the average, but takes into account the size of the bin and thus is statistically more relevant. It represents the number of standard deviations from the average, so centres around 0. The wider the spread, the better the predictor is.

$$\frac{posFraction-negFraction}{\sqrt(\frac{posFraction*(1-posFraction)}{\sum positives}+\frac{negFraction*(1-negFraction)}{\sum negatives})}$$ 

See also: http://techdocs.rpega.com/display/EPZ/2019/06/21/Z-ratio+calculation+in+ADM.

```{r}
binningDerived[, posFraction := Positives/sum(Positives)]
binningDerived[, negFraction := Negatives/sum(Negatives)]
binningDerived[, `Z-Ratio` := (posFraction-negFraction)/sqrt(posFraction*(1-posFraction)/sum(Positives) + negFraction*(1-negFraction)/sum(Negatives))]
```

```{r, echo=F}
binningDerived[, Lift := NULL]

kable(binningDerived) %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  column_spec(c(2,3), bold = T, border_left = T, border_right = T) %>%
    column_spec(6, color = "darkblue") 
```

# Predictor AUC

The predictor AUC is can be derived from the positives and negatives as well, e.g. using the *pROC* package.

```{r, warning=F,echo=F,message=F}
library(pROC)

response = unlist(sapply(1:nrow(predictorbinning),function(r){return(c(rep(T, predictorbinning$Positives[r]), rep(F, predictorbinning$Negatives[r])))}))
prediction = unlist(sapply(1:nrow(predictorbinning),function(r){return(rep(predictorbinning$`Propensity (%)`[r], predictorbinning$Positives[r]+predictorbinning$Negatives[r]))}))
plot.roc(response, prediction, print.auc=T, col="darkgreen", levels=c(T,F), direction=">")
```

The AUC can also be calculated directly from the positives and negatives using a utility function in *cdhtools*:

```{r}
cdhtools::auc_from_bincounts(predictorbinning$Positives, predictorbinning$Negatives)
```

# Predictor score and log odds

The score is calculated from the log odds





