---
title: "ih-reporting"
author: "Pega"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ih-reporting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
# knitr options: https://yihui.name/knitr/options/
knitr::opts_chunk$set(
collapse = TRUE,
fig.width = 7,
fig.height = 5,
fig.align = "center",
comment = "#>"
)
```

# IH Reports

A few simple reports on IH.

Input is a small IH export included in pdstools, you can point to any Pega
export and read with readDSExport.

# Data Loading

```{r message=FALSE, warning=FALSE}
library(pdstools)
library(ggplot2)
library(scales)
library(tidyverse)
library(data.table)
library(colorspace)
library(lubridate)

theme_set(theme_minimal())

data(ihsampledata) 
#ihsampledata <- readDSExport("Data-pxStrategyResult_pxInteractionHistory", "~/Downloads"

pdstools::standardizeFieldCasing(ihsampledata)

# Both "Accept" and "Accepted" occur as labels - fix that for the reporting
ihsampledata[Outcome == "Accept", Outcome := "Accepted"]
```

# Percentage of responses by Channel

This shows how the distribution of responses is different per channel.

```{r, warning=F, message=FALSE}
ggplot(ihsampledata[Channel != ""], aes(Channel, fill=Outcome)) + 
  geom_bar(position = position_fill()) +
  theme_bw() +
  scale_fill_discrete_divergingx() +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() + ylab("") +
  ggtitle("Interaction History", subtitle = "by Channel") + xlab("Channel")
```
# Number of interactions over time

This shows the total number of interactions split by type of response, per
day. Note the additional outcomes "Churned" and "Loyal" here, which are
independent of Channel and have been the result of responses to Customer level
models, not to channel specific propensity models.

```{r message=FALSE, warning=FALSE}
ihsampledata[, Day := as_date(fromPRPCDateTime(OutcomeTime))]
ihsampledata_by_date <- ihsampledata[, .N, by=c("Outcome", "Day")]

ggplot(ihsampledata_by_date, aes(Day, N*1000, fill=Outcome)) +
  geom_area() +
  theme_bw() +
  scale_fill_discrete_divergingx() +
  ggtitle("Interactions over Time") + xlab("Date") + ylab("Number of Interactions")
```

# Calculation of AUC from Interaction History 

The actual model performance can easily be calculated from the data in IH, since
we have both the propensities (from the models) and the actual outcomes.

```{r}
ihsampledata[, Week := week(fromPRPCDateTime(OutcomeTime))]
performance <- ihsampledata[Outcome %in% c("Accept", "Accepted", "Rejected"), 
                  list(AUC = 100*auc_from_probs(Outcome %in% c("Accept", "Accepted"), Propensity),
                    n = .N,
                    Day2 = first(Day)), 
                  by=c("Day", "Name", "Group", "Issue")]

ggplot(performance, aes(Day, AUC, color=Name)) +
  # geom_smooth(size=2,se=F,method = "loess", span=0.5) + 
  geom_line(size=2) + 
  geom_point(size=2) + theme_minimal() +
  scale_color_discrete_qualitative() +
  scale_y_continuous(limits=c(50,NA)) +
  ggtitle("Model Performance", subtitle = "per week")
```


