---
title: "AGB Visualisation"
author: "Pega CDH Tools"
output:
  html_document:
    df_print: paged
    toc: yes
  pdf_document: default
always_allow_html: true
params:
  modeldownload:
    value: "../../data/agb/IHmodelbeforeRetrain.txt"
    # value: "../../data/agb/ModelExportExample.json"
---

```{r setup, include=FALSE}
library(ggplot2)
library(colorspace)
library(data.table)
library(jsonlite)
library(kableExtra)

theme_set(theme_minimal())
options(digits = 5)
knitr::opts_chunk$set(
  comment = ">", echo = FALSE, warning = FALSE
)
```

# AGB Model Visualization

Viz of an export of an AGB model. Should be similar to what you see in 
Pred Studio but with a few more details. Feature importance should
be identical.

This provides a view of just a single model at a given moment in time. Similar
data on multiple models and over time is available in the datamart, in the
model snapshots stored there.

# Read model export

Read the JSON model from an ADM AGB tree exported from Pega.

In early versions (prior to 8.7.3) the model export was a mix of text
and JSON. Later versions have proper JSON. When reading in we try to be 
careful and skip over the non-JSON if it seems to be an old version.

```{r}
linez <- readLines(params$modeldownload)
if (linez[1] != "{") {
  # Old format
  
  modelStart <- which(startsWith(linez, "model=AdaptiveBoostScoringModel{"))
  jsonStart <- which(startsWith(linez, "model={"))

  meta <- strsplit(linez[(modelStart+1):(jsonStart-1)], "=", fixed=T)
  modelJSONText <- paste("{", paste(linez[(jsonStart+1) : (length(linez)-1) ], collapse = "\n"), "}")
  modelJSON <- fromJSON(modelJSONText, flatten = F, simplifyVector = F, simplifyDataFrame = F, simplifyMatrix = F)
  
  modelMetaInfo <- data.table( Attribute = sapply(meta, function(x) x[[1]]), 
                             Value = sapply(meta, function(x) paste(x[2:length(x)], collapse=" = ")))

} else {
  # New all JSON format
  
  allJSON <- fromJSON(paste(linez, collapse = "\n"), flatten = F, simplifyVector = F, simplifyDataFrame = F, simplifyMatrix = F)
  modelJSON <- allJSON$model
  
  modelMetaInfo <- data.table(Attribute = setdiff(names(allJSON), "model"))
  modelMetaInfo[ , Value := toJSON(allJSON[[Attribute]]), by=Attribute]
}
```


# Meta data about the model

```{r}
modelMetaInfo %>% kable() %>% kable_classic()
```

# Model Trees

The model consists of a number of trees. They are all under a single 
booster node (currently).

```{r include=FALSE}
forest <- modelJSON$boosters[[1]]$trees
```

Per tree we are interested in the max depth, the number of nodes,
number of leaf nodes and the gain.

```{r}
maxDepth <- function(tree) {
  if (!("split" %in% names(tree))) return(0)
  
  return(1+max(maxDepth(tree$left), maxDepth(tree$right)))
}

numberOfNodes <- function(tree) {
  if (!("split" %in% names(tree))) return(1)
  
  return(1+numberOfNodes(tree$left)+numberOfNodes(tree$right))
}

numberOfLeafNodes <- function(tree) {
  if (!("split" %in% names(tree))) return(1)
  
  return(numberOfLeafNodes(tree$left)+numberOfLeafNodes(tree$right))
}

getSplit <- function(tree)
{
  # This assumes the "human friendly" format as found in the model download,
  # for the data in the model table we may need a slightly different way to 
  # parse, and also look up the index in the predictors (Index) in the JSON
  # file.
  trimws(strsplit(tree$split," in|<|>|<=|>=|=|is ")[[1]][1])  
}

splits <- function(tree) {
  if (!("split" %in% names(tree))) return(NULL)
  
  return(unique(c(getSplit(tree), splits(tree$left), splits(tree$right))))
}

gain <- function(tree) {
  return(tree$gain)
}

tbl <- data.table( tree = seq(length(forest)),
                   maxDepth = sapply(forest, maxDepth),
                   nrNodes = sapply(forest, numberOfNodes),
                   nrLeafNodes = sapply(forest, numberOfLeafNodes),
                   gain = sapply(forest, gain),
                   splits = sapply(forest, function(x) {paste(splits(x), collapse=", ")}))

tbl %>% kable() %>% kable_classic()
```

## Gain over trees

The gain is expected to decrease by tree.

```{r}
ggplot(tbl, aes(tree, gain)) + 
  geom_line(size=1, color="darkgreen") +
  scale_y_continuous(limits = c(0,NA)) +
  ggtitle("Gains by tree")
```

## Max depth of the trees

The max depth does not necessarily decrease over the number of trees, so 
viewing this as a bar chart instead.

```{r}
ggplot(tbl, aes(tree, maxDepth)) + 
  geom_col(aes(fill=-maxDepth)) +
  scale_fill_continuous(guide="none") +
  ggtitle("Max Depth by tree")
```

## Number of nodes


```{r}
ggplot(melt(tbl, id.vars = "tree", measure.vars = c("nrNodes", "nrLeafNodes"), variable.name = "Type", value.name = "Number"), 
       aes(tree, Number, color=Type)) + 
  geom_line() + geom_point() +
  ggtitle("Number of Nodes")
```

# Predictor Importance

```{r}
getGains <- function(tree)
{
  if (!("split" %in% names(tree))) return(data.table())
  
  return(rbindlist(list(data.table(split = getSplit(tree),
                                   gain = tree$gain), 
                        getGains(tree$right), getGains(tree$left)), fill=T))
}

allSplits <- rbindlist(lapply(forest, getGains))
varImp <- allSplits[, list(Importance = mean(gain)), by="split"][order(-Importance)]

varImp[, Importance := 100*Importance/max(Importance)]
varImp[, split := factor(split, levels = rev(varImp$split))]

ggplot(head(varImp,20), aes(Importance, split)) + 
  geom_col(aes(fill=Importance)) +
  scale_fill_continuous(guide="none") +
  ggtitle("AGB Var Importance", subtitle = params$modeldownload) +
  ylab("")
```

