---
title: "Example ADM Predictor Sensitivity Analysis"
author: "Pega"
date: "3/19/2021"
output: html_document
---

# Sensitivity Analysis

This is an example of using `cdhtools` functions to create a predictor
sensitivity plot for ADM models.

To run this example, export your own ADM datamart data
from Pega. These steps are detailed in that same Wiki, see [ADM Datamart](https://github.com/pegasystems/cdh-datascientist-tools/wiki/CDH-Model-analysis-from-ADM-Datamart).

The `admVarImp` function is also illustrated in one of the vignettes, use
`browseVignettes("cdhtools")`.

## Libraries

Load the necessary libraries

```{r}
library(cdhtools)
library(data.table)
library(ggplot2)
library(colorspace)
theme_set(theme_minimal())
```

## Data

Pull in data from the ADM datamart. For the predictor data we need the detailed
bin level data, hence the flag to `readADMDatamartPredictorExport`. When using
the generic `readDSExport` this is (obviously) not needed.

```{r}
admdatamart_models <- readADMDatamartModelExport("~/Downloads")
admdatamart_binning <- readADMDatamartPredictorExport("~/Downloads", noBinning = F)
```

## Variable Importance

This gives the variable importance across all instances of one of the ADM rules. Replace the
model name ("OmniAdaptiveModel") used here by one that exists in your data
export.

```{r}
varimp <- admVarImp(admdatamart_models[ConfigurationName=="OmniAdaptiveModel"], admdatamart_binning)

varimp[]
```


## Plots

### Global overview

There is an additional method that turns the variable importance data into
a `ggplot` plot.

```{r}
admVarImpPlot(varimp)
```

### Multiple configurations

Split by configuration (Pega rule) name.

Note how we add some decoration to the returned plot object.

```{r}
varimp <- admVarImp(admdatamart_models, admdatamart_binning, facets = "ConfigurationName")

admVarImpPlot(varimp) +
  scale_fill_continuous_diverging(palette="Tropic")
```

### By Issue and Channel

Selecting only one particular rule and spitting by Issue and Channel.

Some more decoration is added to the plot.

```{r}
varimp <- admVarImp(admdatamart_models[ConfigurationName=="OmniAdaptiveModel"], admdatamart_binning, 
                    facets = c("Issue", "Channel"))

admVarImpPlot(varimp) +
  scale_fill_continuous_diverging(palette="Tropic") +
  theme(strip.background = element_rect(color="darkgrey", fill="grey", size=1.5, linetype="solid")
  )
```

