---
title: "Example analysis of pxDecisionResults"
author: "Pega Data Scientist Tools"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
  pdf_document: default
always_allow_html: true
params:
  source:
    # Full path to pxDR dataset export zip
    value: "../../data/pxdr/Data-Decision-Results_pxDecisionResults_20220112T162055_GMT.zip"
    
    # ~/Downloads/Data-Decision-Results_pxDecisionResults_20220112T162055_GMT.zip"

  topNpropositions:
    value: 40
    
  topNdecisions:
    value: 15
---

```{r, echo=F, warning=F, error=F, include=FALSE}
library(pdstools)

# include below when developing the library
# sapply(list.files("~/Documents/pega/pega-datascientist-tools/r/R", "*.R", full.names = T), source)

library(data.table)
library(lubridate)
library(ggplot2)
library(colorspace)
library(scales)
library(knitr)
library(kableExtra)
library(bit64)

theme_set(theme_light())
```

# Intro

The **pxDecisionResults** dataset represents the internal store of the decision
results produced by strategies. They are stored in this dataset if the 
execution mode in the dataflow specifies to keep the decision results 
"for later response capture". This is the standard operating mode.

This dataset can be exported for detailed analysis. Many reports in the 
product are also based on the same data. VBD is one reporting and 
aggregating mechanism on top of it. The decision results are typically also 
stored in IH and - indirectly - in the IH summaries.

This notebook just shows the basics of importing this data and doing a
(trivial) analysis. Such detailed analysis is occasionally needed for 
deeper analysis.

Note that pxDecisionResults is typically very large so often not (easily)
exportable. For analysis purposes you would typically first truncate
it, run a data flow for just a limited number of interactions, then 
export the pxDecisionResults dataset (from Dev Studio).

# Load data

Read the data. pxDR is a nested structure so making sure to include embedded
complex types. The **pxUpdateDateTime** field is a large integer - so making
sure the representation in R is correctly set as 64 bit integer.

```{r}
pxDR <- readDSExport(params$source, excludeComplexTypes = F)
pxDR[["pxUpdateDateTime"]] <- as.integer64(pxDR$pxUpdateDateTime)
dim(pxDR)
```

# Fields

Showing all available fields. Note that the set of fields available depends
on the application layer. Some are from the base DSM layer, some from the
CDH layer, and others from the application layer on top of that.

```{r}
decisionResults <- rbindlist(pxDR$pxDecisionResults, fill = T)
setcolorder(decisionResults, sort(names(decisionResults)))

str(decisionResults)
```


# Previewing the first `r params$params$topNdecisions` decision results

```{r}
decisionResults[order(pxDecisionTime)] %>% head(params$topNdecisions) %>% kbl() %>% kable_paper("hover", full_width = T)
```

# Proposition Distribution

Showing the tope `r params$topNpropositions` propositions.

```{r}
propositionDistributionSummary <- 
  decisionResults[, .(N=.N), by=c("pyName", "pyChannel")]
propositionDistributionSummary[, Total := sum(N), by=pyName]
propositionDistributionSummary[, Rank := rank(Total)]
setorder(propositionDistributionSummary, -Rank)
propositionDistributionSummary[, pyName := factor(pyName, unique(pyName))]

ggplot(propositionDistributionSummary[as.integer(pyName)<=params$topNpropositions], aes(pyName, N, fill=pyChannel)) + geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0)) +
  scale_fill_discrete_qualitative() +
  ggtitle("Number of Propositions", subtitle = paste("top", params$topNpropositions))
```
# Propensity ranges

```{r}
ggplot(decisionResults, aes(pyPropensity, pyChannel, color=pyChannel)) +
  geom_boxplot()
```


