---
title: "Analysis of Model Executions in pxDecisionResults"
author: "Pega CDH Tools"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
  pdf_document: default
always_allow_html: true
params:
  source:
    # Full path to pxDR dataset export zip
    # value: "~/Downloads/Data-Decision-Results_pxDecisionResults_20220112T162055_GMT.zip"
    # value: "~/Downloads/Data-Decision-Results_pxDecisionResults_20220117T123506_GMT.zip"
    # value: "~/Downloads/Data-Decision-Results_pxDecisionResults_20220117T134919_GMT.zip"
    
    # baseline, before bug fix
    # value: "~/Downloads/Data-Decision-Results_pxDecisionResults_20220128T162345_GMT.zip"

    # with bug fixes
    # value: "~/Downloads/Data-Decision-Results_pxDecisionResults_20220128T163121_GMT.zip"
    # value: "~/Downloads/Data-Decision-Results_pxDecisionResults_20220201T085415_GMT.zip"
    
    value: "../../data/pxdr/Data-Decision-Results_pxDecisionResults_20220112T162055_GMT.zip"

  topNinteractions:
    value: 10

  topNdecisions:
    value: 15
    
  indexOfSelectedInteraction:
    value: 1

---

```{r, echo=F, warning=F, error=F, include=FALSE}
library(cdhtools)

# include below when developing the library
# sapply(list.files("~/Documents/pega/cdh-datascientist-tools/r/R", "*.R", full.names = T), source)

library(data.table)
library(lubridate)
library(ggplot2)
library(colorspace)
library(scales)
library(knitr)
library(kableExtra)
library(bit64)
library(plotly)

theme_set(theme_light())
```

# Intro

Detailed analysis of the **pxDR** decision results with a focus on the 
"Model Execution Results". Strategy Results are the outputs of 
a strategy. 

For analysis purposes you would typically first truncate
the **pxDecisionResults** dataset, then run a data flow for just a limited 
number of interactions, then export the dataset.

# Load data

Read the data. pxDR is a nested structure so making sure to include embedded
complex types.

```{r}
pxDR <- copy(readDSExport(params$source, excludeComplexTypes = F))
pxDR[, pxUpdateDateTime := as.integer64(pxUpdateDateTime)]

dim(pxDR)
```

# Number of Model Excecution Results 

## Preview of the first `r params$params$topNdecisions` decision results

```{r include=FALSE}
splitModelExecutionReferences <- function(mrefs)
{
  sapply(strsplit(mrefs,",",fixed=T),as.integer)  
}

countStringModelExecutionReferences <- function(mrefs)
{
  ifelse(sapply(mrefs, is.na), 0L,
         sapply(strsplit(mrefs,",",fixed=T),length))  
}
```

```{r}
decisionResults <- rbindlist(pxDR$pxDecisionResults, 
                             fill = T)[, c("pyOutcome", "pyPropensity", "pyName", "pyChannel", 
                                           "pzModelExecutionReferences", "pyModelPropensity", 
                                           "pxDecisionTime", "pxInteractionID")]

decisionResults[, pxDRindex := unlist(sapply(seq(nrow(pxDR)), function(i){rep(i, nrow(pxDR$pxDecisionResults[[i]]))}))]
decisionResults[, nModelExecutionResults := countStringModelExecutionReferences(pzModelExecutionReferences) ]
setorder(decisionResults, pxDecisionTime)

decisionResults %>% 
  head(params$topNdecisions) %>% kbl() %>% kable_paper("hover", full_width = T)
```

## Totals across all `r length(pxDR)` interactions

```{r}
pxDR[, nDecisionResults := sapply(pxDecisionResults, nrow)]
pxDR[, nModelExecutionResults := sapply(pxModelExecutionResults, nrow)]

data.table( `Number of Interactions` = nrow(pxDR),
            `Number of DecisionResults` = sum(pxDR$nDecisionResults),
            `Number of ModelExecutionResults` = sum(pxDR$nModelExecutionResults)) %>% 
  kbl() %>% kable_paper("hover", full_width = T)
```

## Totals per interaction for the first `r params$topNinteractions` interactions

Showing the first couple of records along with the number of decision 
results and the total number of unique model execution results.

```{r}
pxDR[, c("pxVersion", "pxInteractionID", "pxSubjectID", "nDecisionResults", "nModelExecutionResults", "pxUpdateDateTime")][order(pxUpdateDateTime)] %>% head(params$topNinteractions) %>% 
  kbl() %>% kable_paper("hover", full_width = T)
```

## Decision Results split by Outcome for all decisions

Decision results and model executions split by outcome.

Strategy results typically have only a few model executions, depending on 
the configuration of course. A single model has one, a simple prediction two,
a two-stage prediction 5 and if there are additional models involved in 
evaluating a prediction, more.

```{r}
plt <- ggplot(decisionResults, aes(as.factor(nModelExecutionResults), fill=pyOutcome)) + 
  geom_bar() +
  ggtitle("Number of Decision Results per nr of Model Execution Results", subtitle = "split by Outcome") +
  scale_fill_discrete_qualitative() +
  xlab("Model Execution Results") +
  ylab("# Decision Results") +
  scale_y_continuous(limits=c(0,nrow(decisionResults))) # scale to max to make it easier to compare plots from different runs

plotly::ggplotly(plt)
```

## Decision Results by Channel and Outcome

Showing the number of decision results and models executed for 
each channel and outcome combination.

```{r}
plt <- ggplot(decisionResults[, .(N = .N, nModelExecutionResults = sum(nModelExecutionResults)), 
                              by=c("pyChannel", "pyOutcome")], 
              aes(pyOutcome, pyChannel, label=paste(N, nModelExecutionResults, sep="\n"))) + 
  geom_raster(aes(fill = pyOutcome, alpha = N)) +
  geom_text(size=3) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Decision results by Channel and Outcome", 
          subtitle = "showing # of decision results and # of model executions")

plotly::ggplotly(plt)
```

# Details of the Model Execution Results

ADM models, Predictive Models and Predictions all have different details
in the model execution results.

## Decision Results split by Type of Model execution

The summary below shows the number of decision results split by type of
model execution, and per outcome.

```{r include=FALSE}
modelExecutionDetailsFromJSON <- function(modelExecutionJSON) 
{
  if (is.na(modelExecutionJSON) || length(modelExecutionJSON)==0) {
    return(data.table(modelType = NA,
                      component = NA,
                      positives = NA,
                      negatives = NA,
                      rule = NA))
  }
  mer_list <- jsonlite::fromJSON(modelExecutionJSON)
  return(data.table(modelType = mer_list[["modelType"]],
                    component = mer_list[["component"]],
                    positives = paste(unlist(mer_list[["positiveOutcomes"]]), collapse = "/"),
                    negatives = paste(unlist(mer_list[["negativeOutcomes"]]), collapse = "/"),
                    rule = paste(mer_list[["ruleClass"]], mer_list[["ruleName"]], sep = ":")))  
}
```

```{r}
modelExecutionSummary <- 
  rbindlist(lapply(seq(nrow(decisionResults)),
                   function(i) {
                     merJSONs <- as.data.table(pxDR[decisionResults[i, pxDRindex]]$pxModelExecutionResults[[1]])
                     
                     results <- rbindlist(lapply(merJSONs[splitModelExecutionReferences(decisionResults[i, pzModelExecutionReferences]), 
                                                          pxModelExecutionResults], 
                                                 modelExecutionDetailsFromJSON), 
                                          use.names = T, fill = T)
                     
                     results[, pyOutcome := decisionResults[i, pyOutcome]]
                     
                     results
                   } ), use.names = T, fill = T)

dcast(modelExecutionSummary, modelType+component+positives+negatives+rule ~ pyOutcome, fun.aggregate = length) %>%
  kbl() %>% kable_paper("hover", full_width = T)
```

## Raw JSON of the Model Execution Results of one decision

In the evaluation of every decision a number of models / predictions can 
be hit / executed.

Here we pick out one decision (every interaction has a number of decision results,
and every decision result has a number of model executions) and show the
raw JSON content of these.

The interaction is specified by a parameter as an index.

```{r}
interactionOfInterest <- pxDR[params$indexOfSelectedInteraction]$pxInteractionID

interactionOfInterest
```

For the decision result of that interaction we pick the first one with the
largest number of decision results.

```{r}
srOfInterest <- decisionResults[interactionOfInterest == pxInteractionID][which.max(decisionResults[interactionOfInterest == pxInteractionID, 
                                                                                                    nModelExecutionResults])]

srOfInterest$pzModelExecutionReferences
```

```{r}
rawJSON <- as.data.table(pxDR[srOfInterest$pxDRindex]$pxModelExecutionResults[[1]])

dummy <- sapply(rawJSON[as.integer(splitModelExecutionReferences(srOfInterest$pzModelExecutionReferences)), 
                        pxModelExecutionResults], 
                function(x) { 
                  cat(fill=T)
                  cat(jsonlite::minify(x), fill=T)
                })
```


