---
title: "Hello CDH Tools"
author: "Pega"
date: "1/26/2021"
output: html_document
---

# Hello CDH Tools

This is a basic example of using CDH tools to load data from the ADM datamart
and create some basic visualizations.

To run this first install the `cdhtool` library as per instructions from the 
GitHub repository https://github.com/pegasystems/cdh-datascientist-tools. Follow
the link to the Wiki from the bottom of the page and find installation 
instructions there.

```{r message=FALSE, warning=FALSE}
library(cdhtools)
library(data.table)
```

Then to run this example, export the dataset representing the ADM models from
Pega. These steps are detailed in that same Wiki, see [ADM Datamart](https://github.com/pegasystems/cdh-datascientist-tools/wiki/CDH-Model-analysis-from-ADM-Datamart).

Once you have exported and downloaded the data for the dataset
`Data-Decision-ADM-ModelSnapshot` continue the notebook from here.

Change the folder if it is different from the Mac default location used 
below. The function `readADMDatamartModelExport` will take the latest 
downloaded model export zip from that location. If you need more control
over which file is taken exactly you can also specify the exact file name
or use the generic `readDSExport` function. Type `?readADMDatamartModelExport`
for help or use the links in the Wiki.

```{r}
models <- readADMDatamartModelExport("~/Downloads")
```

This `data.table` contains all the model-level attributes. The Pega `py`
prefixes have been stripped off and casing has been normalized, so you will
find attributes like `ActivePredictors`, not `pyactivepredictors`. 

This data can then easily be used for plotting. Below a few examples. For 
more please see the [Graph Gallery](https://github.com/pegasystems/cdh-datascientist-tools/wiki/CDH-Graph-Gallery) in the Wiki.

## Bubble Chart

For example, a Bubble plot with performance vs. success rate split by issue and channel:

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
library(colorspace)

ggplot(models, aes(Performance, Positives/ResponseCount, color=log(Positives), size=ResponseCount)) +
  geom_point(alpha=0.8) +
  facet_grid(Channel~Issue) +
  scale_size_continuous(guide=NULL) +
  scale_color_continuous_sequential(guide=NULL) +
  scale_y_continuous(limits = c(0, NA), labels = scales::percent) +
  labs(title="Performance vs Success Rate",subtitle = "By Channel and Issue", y="Success Rate")

```

## Next steps

For more examples and sample code please see the [CDH Tools Wiki](https://github.com/pegasystems/cdh-datascientist-tools/wiki).






