---
title: "Hello Pega Data Scientist Tools"
author: "Pega Data Scientist Tools"
date: "`r Sys.Date()`"
output: html_document
---

# Hello CDH Tools

This is a basic example of using CDH tools to visualize the ADM datamart.

To run this first install the `cdhtool` library as per instructions from the 
[Wiki](https://github.com/pegasystems/pega-datascientist-tools/wiki) of the GitHub repository https://github.com/pegasystems/pega-datascientist-tools.

```{r message=FALSE, warning=FALSE}
library(pdstools)
library(data.table)
library(ggplot2)
```

Then to run this example, use a sample dataset included or export your own 
from Pega. These steps are detailed in that same Wiki, see [ADM Datamart](https://github.com/pegasystems/pega-datascientist-tools/wiki/CDH-Model-analysis-from-ADM-Datamart).

Change the folder if it is different from the Mac default location used 
below. The function `ADMDatamart` will read ADM datamart and standardize the
fields returned. It is very flexible and can read from datamart export zips,
from CSV, parquet etc and has a number of arguments to fine tune the behavior.

See `?ADMDatamart` for help or use the links in the Wiki.

```{r}
# To use the sample dataset:
# data(adm_datamart)
# datamart <- adm_datamart

datamart <- ADMDatamart("../../data")

#datamart <- ADMDatamart("~/Downloads")
```

The returned structured contains both the model and predictor data from the
ADM datamart. The Pega `py` prefixes have been stripped off and casing has been 
normalized, so you will find attributes like `ResponseCount`, not `pyResponseCount`. 

This data can then easily be used for plotting. Below a few examples. For 
more please see the [Graph Gallery](https://github.com/pegasystems/pega-datascientist-tools/wiki/CDH-Graph-Gallery) in the Wiki.

## Bubble Chart

For example, a Bubble plot with performance vs. success rate split by issue and channel:

```{r message=FALSE, warning=FALSE}
plotPerformanceSuccessRateBubbleChart(datamart, 
                                      filter = filterLatestSnapshotOnly,
                                      facets=c("Issue", "Channel")) +
  theme_bw()
```

## Predictor Analysis

A quick look at the predictors is just as easy.

The returned plot objects are simply ggplot objects, so additional decoration
can be added.

```{r message=FALSE, warning=FALSE}

plotPredictorImportance(datamart, limit=20) +
  theme_bw()
```

## Next steps

For more examples and sample code see the [Pega Data Scientist Tools Wiki](https://github.com/pegasystems/pega-datascientist-tools/wiki).






