---
title: "Control Group analysis on IH data"
author: "Pega Data Scientist Tools"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
params:
  outputfile:
    value: "./ModelsWithStatSignificantDiffFromRandom.csv"
---

```{r message=FALSE, warning=FALSE}
library(pdstools)
library(data.table)
library(ggplot2)
library(scales)
library(colorspace)
library(lubridate)
library(kableExtra)
# library(tidyverse)

theme_set(theme_minimal())
```

# Intro

Plot Control and Test groups conversion for the action/treatment combination.

NBAD by default uses OOTB 2% of customers for random control group. This
notebook calculates the z-scores and visualizes the significance based on
Accept and Reject outcomes.

For CTR use Clicked and Impression outcomes for Web. Set Total = Impression count,
do not sum with Clicked outcomes

# Data

By default uses the small IH sample that is part of the Pega Data Scientist package.

To use your own data use use the *readDSExport* function with the json file
as argument, or use *read_json_arrow* from the *arrow* package.

Selects Web channel outcomes and set ActionID = Adaptive model identifier if using treatment level models
Treatment level models is a preferred approach for NBAD based CDH implementations.

Use Issue/Group/Name if using action level models

```{r}
data("ihsampledata")

# standardized camel casing of fields
standardizeFieldCasing(ihsampledata)

# Both "Accept" and "Accepted" occur as labels - fix that for the reporting
ihsampledata[Outcome == "Accept", Outcome := "Accepted"]

ih <- ihsampledata[(Outcome == "Accepted" | Outcome == "Rejected") & (Channel == "Web")]
ih[, ActionID := paste(Issue, Group, Name, Treatment, sep = "/")]

# Filter IH data by action id filter: may be Issue/Group/Name/Treatment
action_id_filter <- ""

```

# Calculations

Function to calculate Z-score, p-value and confidence interval, as soon as R
performs Chi^2 test should be used if control group size > 30.

```{r}
z.prop <-
  function(x1,
           x2,
           n1,
           n2,
           conf.level = 0.95,
           alternative = "two.sided") {
    numerator <- (x1 / n1) - (x2 / n2)
    p.common <- (x1 + x2) / (n1 + n2)
    denominator <-
      sqrt(p.common * (1 - p.common) * (1 / n1 + 1 / n2))
    z.prop.ris <-
      numerator / denominator
    p1 <- x1 / n1
    p2 <- x2 / n2
    p.val <- 2 * pnorm(-abs(z.prop.ris))
    if (alternative == "lower") {
      p.val <- pnorm(z.prop.ris)
    }
    if (alternative == "greater") {
      p.val <- 1 - (pnorm(z.prop.ris))
    }
    SE.diff <- sqrt(((p1 * (1 - p1) / n1) + (p2 * (1 - p2) / n2)))
    diff.prop <- (p1 - p2)
    CI <- (p1 - p2) + c(-1 * ((qnorm(((1 - conf.level) / 2
    ) + conf.level)) * SE.diff),
    ((qnorm(((1 - conf.level) / 2
    ) + conf.level)) * SE.diff))

    return (list(
      estimate = c(p1, p2),
      ts.z = z.prop.ris,
      p.val = p.val,
      diff.prop = diff.prop,
      CI = CI
    ))

  }

#Substring from the nth character
substrLeft <- function(x, n) {
  substr(x, n + 1, nchar(x))
}
```


Calculate acceptance rate/click-through rate

```{r}
ih_analysis <- dcast(ih[Modelcontrolgroup != "", list(Count = .N), 
                        by=c("ActionID", "Modelcontrolgroup", "Outcome")],
      ActionID+Modelcontrolgroup ~ Outcome, value.var = "Count")
ih_analysis[, Total := Accepted + Rejected]
ih_analysis[, Conversion := 100 * Accepted / Total]
ih_analysis[, ModelName := gsub(".*/([^/]*)/","\\1",ActionID)]

#Filter out actions
ih_action <- ih_analysis[startsWith(ActionID, action_id_filter)]
```


# Plot acceptance rate for test and control groups

```{r}
hist_action_cmp <- ih_action %>%
  ggplot(aes(x = Modelcontrolgroup, y = Conversion)) +
  geom_col(aes(fill = Modelcontrolgroup),
           color = "darkslategray",
           alpha = 0.5) +
  facet_grid(
    ~ gsub("/", "\n", ModelName),
    scales = "free_x",
    space = "free_x",
    switch = "x"
  ) +
  ggtitle("Control group conversion analysis") +
  labs(y = "Accept Rate %", x = "Action/Treatment model", fill = "Group")

hist_action_cmp
```


# Calculate Z-score and p-value

```{r}
evaluate_models <- dcast(ih_action, ActionID+ModelName ~ Modelcontrolgroup, value.var=c("Accepted","Rejected","Total","Conversion"))
evaluate_models[, `Z Score` := z.prop(Accepted_Control, Accepted_Test, Total_Control, Total_Test)$ts.z]
evaluate_models[, `P value` := z.prop(Accepted_Control, Accepted_Test, Total_Control, Total_Test)$p.val]

z_test <- evaluate_models %>%
  ggplot(aes(x = gsub("/", "\n", ModelName), y = `Z Score`, fill = `Z Score`)) +
  geom_col() +
  ggtitle("Z-score") +
  labs(y = "Z", x = "Action/Treatment") +
  scale_fill_continuous(guide="none")

z_test
```

# Keep table with models sorted by p-value. 

Saving the table to a file for external use.

```{r}
readr::write_excel_csv(evaluate_models, params$outputfile)

evaluate_models %>% kbl() %>% kable_paper("hover", full_width = F)
```


